Experiences in Modelling a Microcontroller Instruction Set Using B

### 2012-01-20 trunk/papers/ArticleZ80/source/paper.tex
Mais alguns ajustes realizados

\usepackage[english]{babel} %\usepackage{supertabular} % used in B Symbols
\usepackage{hyperref} % add URL in bibliography
\usepackage{url} % allow add URL in document

@@

In general, the manuals of microcontrollers can be more suited for \marginpar{DD: o título da sub-seção não condiz com seu conteúdo. Não cita nenhum problema na formalização. Não vejo muito bem o que ele traz de novo que já não foi dito anteriormente.}
developers in assembly language. Many times, each instruction of
microcontrollers is shown in its official manual using textual
description, math description, examples, encoding, number of cycles
and other information. Sometimes, the instruction descriptions have
an own notation, some informations are organized in different sections
of document and the textual description are not standardized. This
does not allow the interpretation by parser.  The manuals also have
semi-formal and informal descriptions, these descriptions may make
mistakes.  The official manual of the microcontroller Z80 has also
several problems that were identified over
time~\cite{UndocumentedZ80}.

@@

	\bf SETS\\
	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0 \rm , \it f0 \rm ,\it b0 \rm , \it c0 \rm ,\it d0 \rm , \it e0 \rm ,\it h0 \rm , \it l0 \rm ,\it a\_0 \rm , \it f\_0 \rm ,\it b\_0 \rm , \it c\_0 \rm ,\it d\_0 \rm , \it e\_0 \rm ,\it h\_0 \rm , \it l\_0 \rm \};\\
	\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC \rm , \it DE \rm , \it HL \rm , \it SP \rm , \it AF \rm \}

@@

\textbf{Assuring the absence of overlapping of address regions:}
 \emph{To assure that address regions are well defined, then the designer must to verify the expression:}
%\begin{sloppypar}
%\hspace*{0.10in}
\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ STACK\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ PROGRAM\_R\_ADR $\cap$  STACK\_R\_ADR  $=$ \{\} 
%\end{sloppypar}

\textbf{Preserving the consistency of the memory:} \emph{In general, the access to
some address regions is dangerous. Then, each instruction has a specific pre-condition that verify
if the new address memory, that will be updated, is member of its region. For example, the $\textit{PUSH}$
program instruction allows write only in the region of stack ($\textit{STACK\_R\_ADR}$).}

@@

\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR

@@

\hspace*{0.20in}\bf ANY

\hspace*{0.40in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry

@@

@@

\section{Related works}
\label{sec:relatedworks}

>>>>

\usepackage[english]{babel} %\usepackage{supertabular} % used in B Symbols
%\usepackage{hyperref} % add URL in bibliography
\usepackage{url} % allow add URL in document

@@

In general, the manuals of microcontrollers can be more suited for \marginpar{DD: o título da sub-seção não condiz com seu conteúdo. Não cita nenhum problema na formalização. Não vejo muito bem o que ele traz de novo que já não foi dito anteriormente.}
developers in assembly language. Many times, each instruction of
microcontrollers is shown in its official manual using textual
description, math description, examples, encoding, number of cycles
and other information. Sometimes, the instruction descriptions have
an own notation, some informations are organized in different sections
of document and the textual description are not standardized. 
The manuals also have semi-formal and informal descriptions, these descriptions
may leave errors, inconsistency and ambiguity.  
For example in official manual of the microcontroller Z80 \cite{Z80_manual}, if the users \marginpar{VM: Adicionei um exemplo real e uma referência  que cita vários problemas identificados no Manual do Z80.} 
want to know the action of an instruction and its effects on the flag register then he needs
to read an informal textual description, that  some times can be placed in different pages. 
Furthermore, the official manual of the microcontroller Z80 has also several
problems that were identified over time~\cite{UndocumentedZ80}.

@@

	\bf SETS\\
	\hspace*{0.10in}\it id\_reg\_8 \rm = \rm \{ \it a0\rm, \it f0\rm,\it b0\rm, \it c0\rm,\it d0\rm, \it e0\rm,\it h0\rm, \it l0\rm,\it a\_0\rm, \it f\_0\rm,\it b\_0\rm, \it c\_0\rm,\it d\_0\rm, \it e\_0\rm,\it h\_0\rm, \it l\_0 \rm \};\\
	\hspace*{0.10in}\it id\_reg\_16 \rm = \rm \{ \it BC\rm, \it DE\rm, \it HL\rm, \it SP\rm, \it AF \rm \}

@@

\textbf{Assuring the absence of overlapping of address regions:}
 \emph{To assure that address regions are well defined, then the designer must to verify the expression:}
%\begin{sloppypar}
%\hspace*{0.10in}
\it PROGRAM\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ STACK\_R\_ADR $\cap$ DATA\_R\_ADR $\cap$ $=$ \{\} $\land$ PROGRAM\_R\_ADR $\cap$  STACK\_R\_ADR  $=$ \{\} \rm
%\end{sloppypar}

\textbf{Preserving the consistency of the memory:} \emph{In general, the access to
some address regions is dangerous. Then, each instruction has a specific pre-condition that verify
if the new address memory, that will be updated, is member of its region. For example, the $\textit{PUSH}$
program instruction allows write only in the region of stack ($\textit{STACK\_R\_ADR}$\rm).}

@@

\hspace*{0.20in}\bf PRE \it nn $\in$ \it USHORT\hspace*{0.15in} $\land$ \hspace*{0.10in}\it nn\hspace*{0.10in} $\in$  \it DATA\_R\_ADR \hspace*{0.10in}\bf THEN

@@

\hspace*{0.20in}\bf ANY \hspace*{0.10in}\it negative \rm , \it zero \rm , \it half\_carry \rm , \it pv \rm , \it add\_sub \rm , \it carry

@@

\hspace*{0.20in}\bf END

@@

\section{Related work}
\label{sec:relatedworks}
